---
title: "HTSSIP: beta diversity ordinations"
author: "Nick Youngblut"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Beta diversity ordinations

## Dataset

First, let's load some packages including \code{HTSSIP}. Also let's get an overview of the phyloseq object that we're going to use.

```{r, message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(HTSSIP)
physeq
```

## Parsing the dataset

The first task is to split apart the dataset so that we can compare
gradient fractions of each labeled-treatment (e.g., 13C-Cellulose) 
to their corresponding unlabeled control in order to identify
buoyant density ranges with high beta-diversity, which would suggest
a BD "shift" of "heavy" DNA in the labeled treatment due to
isotope incorporation.

Now, let's get the parameters that can be used to subset the dataset into
subsets that contain fractions from just one gradient. 

```{r}
params = get_treatment_params(physeq, c('Substrate', 'Day'))
params
```

We just need the parameters for each treatment, so let's filter out the controls.
In this case, the controls are all '12C-Con'. 

```{r}
params = dplyr::filter(params, Substrate!='12C-Con')
```

Now, we will use an expression that will subset the phyloseq object into the comparisons
that we want to make. 

\code{ex} is an expression that will be used for pruning the phyloseq object

```{r}
ex = "(Substrate=='12C-Con' & Day=='${Day}') | (Substrate=='${Substrate}' & Day == '${Day}')"
physeq_l = phyloseq_subset(physeq, params, ex)
physeq_l
```


## Calculating ordinations

Now, let's actually make ordinations of each treatment compared to the control. 

```{r, message=FALSE, warning=FALSE}
physeq_l_p = SIP_betaDiv_ord(physeq_l)
```

We now have a list of ggplot figures:

```{r, fig.height=4, fig.width=6.5, warning=FALSE}
n = names(physeq_l_p)
physeq_l_p[[1]]
physeq_l_p[[2]]
```



# Session info

```{r}
sessionInfo()
```
