---
title: "HTS-SIP data simulation"
author: "Nick Youngblut"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{HTS-SIP data simulation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Introduction

HTSSIP has some functionality for simulating basic HTS-SIP datasets. With that said, I recommend using a more sophisticated simulation toolset such as SIPSim **[LINK]** for applications other than simple testing of HTS-SIP data analysis functions.

HTSSIP relies *heavily* on the great R package [coenocliner](https://cran.r-project.org/web/packages/coenocliner/index.html). See [this tutorial](http://www.fromthebottomoftheheap.net/code/r-packages/coenocliner/) for a short and simple introduction. 

# Simulating a HTS-SIP dataset

In this vignette, we're going to simulate gradient fraction communities for 6 gradients, with the basic experimental setup as follows:

* Treatments: 13C-glucose vs 12C-control 
* Treatment replicates: 3 (each)
* Total gradients: 6
* Fractions per gradient: 24

First, let's load some packages including `HTSSIP`. 

```{r, message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(HTSSIP)
```

OK, let's set the parameters needed for community simulations. We are basically going to follow the coenocliner tutorial, but instead of a transect along an environmental gradient, we are simulating communities in each fraction of a density gradient. 

```{r}
# setting parameters for tests
set.seed(2)                                # reproducible
M = 6                                      # number of OTUs (species)
ming = 1.67                                # gradient minimum...
maxg = 1.78                                # ...and maximum
nfrac = 24                                 # number of gradient fractions
locs = seq(ming, maxg, length=nfrac)       # gradient locations
tol  = rep(0.005, M)                       # species tolerances
h    = ceiling(rlnorm(M, meanlog=11))      # max abundances

opt = rnorm(M, mean=1.71, sd=0.008)        # species optima (drawn from a normal dist.)
params = cbind(opt=opt, tol=tol, h=h)      # put in a matrix
```

With the current parameters, we can simulate the gradient fraction communities for 1 density gradient: 

```{r}
df_OTU = gradient_sim(locs, params)
df_OTU
```


As you can see, the abundance distribution of each OTU is approximately Gaussian, with varying optima among OTUs. 


### Simulating all communities

Let's make the communities of all gradient fractions for all gradients. We could use a bunch of largely redundant, copy-paste code to do this, but instead, let's use plyr::mlply. 

If all OTUs in the 13C-treatment incorporated labeled isotope, then their abundance distributions should be shifted to 'heavier' buoyant densities. Let's set the 13C-treatment gradients to have a higher mean species optima. For kicks, let's also increase the species optima variance (representing more variable isotope incorporation percentages).


```{r}
# mu = mean species optima
# sd = standard deviation of species optima
df = data.frame(
  mu = c(rep(1.7, 3), rep(1.72, 3)),
  sig = c(rep(0.005, 3), rep(0.008, 3))
)

# a function to make the parameters for each gradient 
fun = function(mu, sig, M, tol, h){
  opt = rnorm(M, mean=1.7, sd=0.005)      # species optima 
  params = cbind(opt=opt, tol=tol, h=h)   # matrix of parameters
  return(params)
}

# making a named list of parameters for each gradient
params_all = plyr::mlply(df, fun, M=M, tol=tol, h=h)
names(params_all) =  c('12C-Con_rep1', '12C-Con_rep2', '12C-Con_rep3', 
                       '13C-Glu_rep1', '13C-Glu_rep2', '13C-Glu_rep3')
class(params_all)
```

**Note:** we need a named list of parameters for the next step. The names will be used as sample IDs. 


### Additional sample metadata

```{r}
meta = data.frame(
  'Gradient' = c('12C-Con_rep1', '12C-Con_rep2', '12C-Con_rep3',
                 '13C-Glu_rep1', '13C-Glu_rep2', '13C-Glu_rep3'),
  'Treatment' = c(rep('12C-Con', 3), rep('13C-Glu', 3)),
  'Replicate' = c(1:3, 1:3)
)
```


```{r}
## physeq object
physeq_rep3 = HTSSIP_sim(locs, params_all, meta=meta)
physeq_rep3
```

```{r}
physeq_rep3 %>% sample_data %>% head %>% as.data.frame
```

```{r}
control_mean_fun = function(x) dnorm(x, mean=1.70, sd=0.01) * 1e8
control_sd_fun = function(x) control_mean_fun(x) / 3
treat_mean_fun = function(x) dnorm(x, mean=1.75, sd=0.01) * 1e8
treat_sd_fun = function(x) treat_mean_fun(x) / 3
```



```{r}
physeq_rep3_qPCR = qPCR_sim(physeq_rep3,
                control_expr='Gradient=="12C-Con"',
                control_mean_fun=control_mean_fun,
                control_sd_fun=control_sd_fun,
                treat_mean_fun=treat_mean_fun,
                treat_sd_fun=treat_sd_fun)

physeq_rep3_qPCR
```


```{r}
physeq_rep3_qPCR %>% otu_table .[1:5,1:5]
```
