---
title: "HTSSIP basics"
author: "Nick Youngblut"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

**HTSSIP** is a package for analyzing high throughput sequence data from DNA- or RNA-SIP experiments. 

I recommend starting with this vignette, then checking out the other vignettes:

* **Dataset exploration**
    * [Beta diversity ordinations of gradient fraction communities](beta_diversity_ordinations.html)
    * [Whole-community BD shift plots](BD_shifts.html)
* **Detecting incorporators**
    * [HR-SIP & MW-HR-SIP](MW_HR_SIP.html)
    * [q-SIP](qSIP.html)
* **Calculating amount of incorporation**
    * [delta-BD](delta_BD.html)
    * [q-SIP](qSIP.html)

# HTSSIP data exploration

For this example, we have 3 gradients, each with ~24 samples (gradient fractions).
The treatment gradients contained DNA extracted from microcosms
recieving both cellulose and glucose, but only one substrate was 13C-labeled
for each treatment. For example, the 13C-cellulose (13C-Cel) treatment recieved
13C-cellulose and 12C-glucose. This experimental design had the benefit of 
requiring only one 12C-control that recieved 12C-cellulose and 12C-glucose. 

The dataset is in phyloseq format 
(currently a requirement for using the HTSSIP package). 
See the [GitHub page](https://github.com/joey711/phyloseq) or the 
[package website](http://joey711.github.io/phyloseq/) for more info 
about the phyloseq. 


First, let's load some packages including \code{HTSSIP}. 

```{r, message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(HTSSIP)
```

Now, let's check out the basic HTSSIP phyloseq object included in the package. 

```{r, message=FALSE, warning=FALSE}
physeq
```

Here's a look at (some) of the sample metadata. Note that **'Bouyant_density'** is
included. This is necessary for many of the analyses in HTSSIP!

```{r}
physeq %>% sample_data %>% .[1:5,1:10]
```


# References for methods in HTSSIP

* **HTS-SIP data exploration**
    * Multiple window High Resolution Stable Isotope Probing (MW-HR-SIP) as described in Youngblut and Buckley *in review*

* **HR-SIP**
    * High Resolution Stable Isotope Probing (HR-SIP) as described in [Pepe-Ranney et al., 2016](http://www.ncbi.nlm.nih.gov/pmc/articles/PMC4867679/)

* **MW-HR-SIP**
    * Multiple window High Resolution Stable Isotope Probing (MW-HR-SIP) as described in Youngblut and Buckley *in review*

* **qSIP**
    * Quantitative stable isotope probing (q-SIP) as described in [Hungate et al., 2015](http://www.ncbi.nlm.nih.gov/pmc/articles/PMC4592864/)


# Session info

```{r}
sessionInfo()
```
