---
title: "Quantifying isotope incorporation"
author: "Nick Youngblut"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


# Introduction

# Dataset 

First, let's load some packages including `HTSSIP`. 

```{r, message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(HTSSIP)
```

OK. We're going to be using 2 data files: 

* HTS-SIP data (phyloseq format)
* qPCR data (total 16S rRNA gene copies per gradient fraction)

We'll be using the dataset that we simulated in the [HTSSIP_sim](./HTSSIP_sim.html) vignette.

The phyloseq object is similar to the dataset in the other vignettes.

```{r}
# HTS-SIP data
physeq_rep3
```

The associated qPCR data is, in this case a list of length = 2. 

```{r}
# qPCR data (list object)
physeq_rep3_qPCR %>% names
```

# q-SIP

```{r}
# transforming OTU counts
physeq_rep3_t = OTU_qPCR_trans(physeq_rep3, physeq_rep3_qPCR)

# calculating atom fraction excess
atomX = qSIP_atom_excess(physeq_rep3_t,
                         control_expr='Treatment=="12C-Con"',
                         treatment_rep='Replicate')
atomX %>% names
```


```{r}
# calculating bootstrapped CI values
df_atomX_boot = qSIP_bootstrap(atomX, n_boot=100)
df_atomX_boot %>% head
```

Let's use the same threshold as Hungate et al., (2015) for defining incorporators. 

```{r}
CI_threshold = 0
df_atomX_boot = df_atomX_boot %>%
  mutate(Incorporator = A_CI_low > CI_threshold,
         OTU = reorder(OTU, -A))
```

How many incorporators?

```{r}
n_incorp = df_atomX_boot %>%
  filter(Incorporator == TRUE) %>%
  nrow
cat('Number of incorporators:', n_incorp, '\n')
```

Plot the results!

```{r, fig.height=3, fig.width=7}

ggplot(df_atomX_boot, aes(OTU, A, ymin=A_CI_low, ymax=A_CI_high, color=Incorporator)) +
  geom_point(size=0.25) +
  geom_linerange() +
  geom_hline(yintercept=0, linetype='dashed', alpha=0.5) +
  labs(x='OTU', y='Atom fraction excess') +
  theme_bw() +
  theme(
    axis.text.x = element_blank()
  )
```



# delta_BD

```{r}
df_dBD = delta_BD(physeq_rep3,
               control_expr='Treatment=="12C-Con"')
```

```{r}
df_dBD = df_dBD %>%
  mutate(OTU = reorder(OTU, -delta_BD))
```

```{r, fig.height=3, fig.width=7}
ggplot(df_dBD, aes(OTU, delta_BD)) +
  geom_point(size=0.25) +
  geom_hline(yintercept=0, linetype='dashed', alpha=0.5) +
  labs(x='OTU', y='delta BD') +
  theme_bw() +
  theme(
    axis.text.x = element_blank()
  )
```


# Comparing methods

```{r}
stopifnot(nrow(df_atomX_boot) == nrow(df_dBD))
df_j = dplyr::inner_join(df_atomX_boot, df_dBD, c('OTU'='OTU'))
stopifnot(nrow(df_atomX_boot) == nrow(df_j))

df_j = df_j %>%
  dplyr::mutate(OTU = reorder(OTU, -delta_BD))

# plotting
ggplot(df_j, aes(OTU, Z, color=Incorporator)) +
  geom_point(size=0.5) +
 # geom_linerange() +
  geom_point(aes(y=delta_BD), color='black', size=0.5) +
  geom_hline(yintercept=0, linetype='dashed', alpha=0.5) +
  labs(x='OTU', y='Atom fraction excess') +
  theme_bw() +
  theme(
    axis.text.x = element_blank()
  )
```


# Session info

```{r}
sessionInfo()
```
